---
title: "EPI-590R Final"
author: "Olivia Peterson"
format: html
editor: visual
---

```{r reading in libraries and data}
library(readxl)
library(tidyverse)
library(gtsummary)
library(here)
library(gtsummary)
library(knitr)

WeldingFumes <- read_excel(here::here("Data/WeldingFumes_2005.xlsx"))

BigWeldingFumes <- read_excel(here::here("Data", "WeldingFumes_08.02.2024.xlsx"))
warning()
```

## Table of Descriptive Statistics

```{r}
#| eval:true
#| label: tbl-one
#| tbl-cap: "This is a table of descriptive statistics about the job and age of the employees sampled stratified by the mine status"

table1 <- tbl_summary(
  WeldingFumes,
  by = CURRENT_MINE_STATUS,
  include = c(JOB_CD_DESC, AGE, MONTH)) |> 
  add_overall(last = TRUE)
table1

mean_age <- inline_text(table1, variable = "AGE", column = "stat_0")
```

The median and IQR are `r mean_age.` @tbl-one contains a table of descriptive statistics that are partly shown in a figure in @fig-bar.

## Regression Results

Add description Inline code printing something like we did with the mean age in the example

```{r}
table2 <- tbl_uvregression(
	WeldingFumes,
	y = C_OVER_E,
	include = c(CONTAMINANT_DESC,
							STATE, MONTH, AGE),
							method = lm)

inline_text(table2, variable = "AGE",
						pattern = "{estimate} ({conf.low}, {conf.high})")
```

## Figure

```{r}
#| label: fig-bar
#| fig-cap: "This is a bar graph of the samples collected in 2025 by month"
counts <- table(WeldingFumes$MONTH)
p <- barplot(counts, main="Samples Collected in 2005 by Month",
				xlab="Month sample was taken",
				ylab="Frequency of samples taken")
```

@fig-bar contains a bar graph of some of the information seen in the table @tbl-one.

## Data explanation

A brief description of the data, including its source

## Standard deviation function

```{r}
x <- c(1,3,5,7,9)

newSD <- function(x) {
	n <- length(x)
	SD_val <- sqrt(sum((x-mean(x))^2) / n-1)
	return(SD_val)
		}

newSD(x)

WeldingFumes$EXPOSURE_LIM <- as.numeric(WeldingFumes$EXPOSURE_LIMIT)
newSD(WeldingFumes$EXPOSURE_LIM)
```
