---
title: "EPI-590R Final"
author: "Olivia Peterson"
format: html
editor: visual
---

```{r reading in data, warning=FALSE}
library(readxl)
library(tidyverse)
library(gtsummary)
library(here)
library(gtsummary)
library(knitr)

WeldingFumes <- read_excel(here::here("Data/WeldingFumes_2005.xlsx"))

BigWeldingFumes <- read_excel(here::here("Data", "WeldingFumes_08.02.2024.xlsx"))
```

## Table of Descriptive Statistics

```{r descriptive stat table}
#| label: tbl-one
#| tbl-cap: "This is a table of descriptive statistics about the job and age of the employees sampled stratified by the mine status"
table1 <- tbl_summary(
  WeldingFumes,
  by = CURRENT_MINE_STATUS,
  include = c(JOB_CD_DESC, AGE, MONTH)) |> 
  add_overall(last = TRUE)
table1
mean_age <- inline_text(table1, variable = "AGE", column = "stat_0")
mean_age
```

@tbl-one contains a table of descriptive statistics partically shown in the figure @fig-bar. The median and IQR for the age of workers sampled is `r mean_age`

## Regression Results

```{r regression table}
#| tbl-cap: "This is a regression table showing statistics for the welding fume elements sampled, as well as the age, month, and state of the employees sampled"
table2 <- tbl_uvregression(
	WeldingFumes,
	y = C_OVER_E,
	include = c(CONTAMINANT_DESC,
							STATE, MONTH, AGE),
							method = lm)
table2

age_reg <- inline_text(table2, variable = "AGE",
						pattern = "{estimate} ({conf.low}, {conf.high})")
```

The regression statistics for the age of the employees sampled is `r age_reg`

## Figure

```{r bar chart figure}
#| label: fig-bar
#| fig-cap: "This is a bar graph of the samples collected in 2025 by month"
counts <- table(WeldingFumes$MONTH)
p <- barplot(counts, main="Samples Collected in 2005 by Month",
				xlab="Month sample was taken",
				ylab="Frequency of samples taken")
```

@fig-bar contains a bar graph of some of the information seen in the table @tbl-one. 

## Data explanation

This data is from a project I have been working on this summer with the CDC's occupational health branch, NIOSH. We are trying to examine the health effects of welding fumes in those that mine in metal and nonmetal mines. The data is public-use from MSHA and was made by combining Personal Health data from miners and mine information. I decided to only include observations from the year 2005 so that the data was more manageable with 300 observations as opposed to 40,000 in the entire data set.

## Standard deviation function

```{r new function}
x <- c(1,3,5,7,9)

newSD <- function(x) {
	n <- length(x)
	SD_val <- sqrt(sum((x-mean(x))^2) / n-1)
	return(SD_val)
		}

newSD(x)

WeldingFumes$EXPOSURE_LIM <- as.numeric(WeldingFumes$EXPOSURE_LIMIT)
newSD(WeldingFumes$EXPOSURE_LIM)
```
